Issues with Dynamical Measurements Across the Disk

### 2019-02-25 paper/ms.tex
last changes to section 3 (for now)

\subsection{Local Midplane} \label{ssec:local_midplane}
Using the three simulations, we determine the local midplane that an
observer might measure if they were situated in each of these galaxies, as a function of azimuth at the solar circle. Starting from the coordinate system described in the previous section, which is approximately aligned so that the $z$ coordinate is perpendicular to the disk plane, we
place our imaginary observer at $z=0$ and a solar radius of $8.2\,\kpc$ and vary the azimuth between 
$0<\phi<2\pi$\footnote{Our very long-lived observer also has warp-drive.}. At each value of $\phi$ we then
compute the median $z$ for stars within a cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$
perpendicular to the fiducial disk. We then re-define this median $z$ as the new midplane of the cylinder, re-select stars, and iterate until the median $z$ value converges. We find that only $10$ iterations of this procedure
are necessary for convergence. The resulting median $z$ is taken to be what our observer would measure as the local
Galactic midplane at each $\phi$.
To account for the effect of particle noise, we bootstrap-resample stars within the cylinder $1000$ times and determine
the $95^{\text{th}}$ minus $5^{\text{th}}$ percentile range. This bootstrap resampling is performed by first selecting all stars
within a height of $2\,\kpc$, resampling that selection, and then repeating
the $10$ iterations.

To allow for potential small inaccuracies in the determination of the original fiducial coordinate system,
we also subtract the best fit $\text{A} \sin{\left(\phi + \text{B}\right)} +
\text{C}$ curve from the midplane as a function of azimuth. For $\text{A}$ the
values are $-174$, $62$, and $3.8\,\pc$, for $\text{B}$ the values are $0.3$,
$0.65$, and $0.021\,\text{rad}$ and for $\text{C}$ the values are $-47$, $18$,
and $-15\,\pc$ for \mi , \mf , and \mm , respectively. For the assumed solar
radius of $8.2\,\kpc$, we can approximate the angle offset for the $z$-axis
from the values of $\text{A}$ --- we compute $1.22$, $0.43$, and $0.03\,\deg$
for \mi , \mf , and \mm . These angle offsets are consistent with the values
given in \citet{2018arXiv180610564S} for the difference between the $z$-axis
as defined by the gas and stars.

Figure~\ref{fig:midplane} shows the inferred midplane our imaginary observer
would make as a function of azimuth for each galaxy. The 5--95 percentile range from the bootstrap procedure is shown as the dashed-line error bars. The $90\%$ interquartile
range for each galaxy is $185$, $162$, $84\,\pc$ for \mi , \mf , and \mm .

\begin{figure*}
\plotone{fig/midplane_fit.pdf}
\caption{The local midplane determined at the fiducial Solar radius
($8.2\,\kpc$) for the three FIRE galaxies \mi , \mf , and \mm\ ({\em left},
{\em center}, and {\em right} panels). The local midplane is determined at a
position $\phi$ by taking the median height of all stars within $R=0.5\,\kpc$
and $z=1\,\kpc$ (in cylindrical coordinates). The procedure is performed again
using the new height $10$ times to converge on the local midplane height. In
order to allow for the possibility that the fiducial Galactocentric coordinate
system is incorrect, we subtract the best fit $A\sin{(\phi+B)}+C$ curve from
each panel --- this figure is reproduced with the original midplane
determination (i.e. before subtracting the best fit sine curve) in
Appendix~\Gus{x}. We then bootstrap resample $1000$ times on all stars within
a $2\,\kpc$ height of the fiducial midplane to determine error bars (95\uth
and 5\uth percentiles), which we report as dashed lines.}
\label{fig:midplane}
\end{figure*}

\subsection{LSR Variations} \label{ssec:lsr_var}
For similar reasons to the midplane varying across the disk, we also expect
that the local standard of rest (LSR) is a function of azimuth as well. We
perform this calculation in Appendix~\ref{app:lsr}, but performing a
best-fit subtraction to correct for misalignment of the original coordinate system (as in the previous section) is more involved, and we
delay to future work. We do note, however, that the variation in the LSR is
less pronounced than for the midplane, at least for how we calculate the LSR.

>>>>

\subsection{Local Midplane} \label{ssec:local_midplane}
Using the three simulations, we determine the local midplane that an
observer might measure if they were situated in each of these galaxies, as a function of azimuth at the solar circle. Starting from the coordinate system described in the previous section, which is approximately aligned so that the $z$ coordinate is perpendicular to the disk plane, we
place our imaginary observer at $z=0$ and a solar radius of $8.2\,\kpc$ and vary the azimuth between 
$0<\phi<2\pi$\footnote{Our very long-lived observer also has warp-drive.}. At each value of $\phi$ we then
compute the median $z$ for stars within a cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$
perpendicular to the fiducial disk. We then re-define this median $z$ as the new midplane of the cylinder, re-select stars, and iterate until the median $z$ value converges. We find that only $10$ iterations of this procedure
are necessary for convergence. The resulting median $z$ is taken to be what our observer would measure as the local
Galactic midplane at each $\phi$. This procedure assumes perfect density estimation, and therefore perfect corrections for extinction, within the cylinder defining the ``solar neighborhood.'' Imperfect extinction correction is likely to increase the amplitude of the estimated fluctuations in $z$.
To account for the effect of particle noise, we bootstrap-resample stars within the cylinder $1000$ times and determine
the $95^{\text{th}}$ minus $5^{\text{th}}$ percentile range. This bootstrap resampling is performed by first selecting all stars
within a height of $2\,\kpc$, resampling that selection, and then repeating
the $10$ iterations.

To allow for potential small inaccuracies in the determination of the original fiducial coordinate system,
we also subtract the best fit $\text{A} \sin{\left(\phi + \text{B}\right)} +
\text{C}$ curve from the midplane as a function of azimuth to account for an overall tilt of the midplane (a simplified version of the strategy described in \citealt{2019ApJ...871..145A}). For $\text{A}$ the
values are $-174$, $62$, and $3.8\,\pc$, for $\text{B}$ the values are $0.3$,
$0.65$, and $0.021\,\text{rad}$ and for $\text{C}$ the values are $-47$, $18$,
and $-15\,\pc$ for \mi , \mf , and \mm , respectively. For the assumed solar
radius of $8.2\,\kpc$, we can approximate the angle offset for the $z$-axis
from the values of $\text{A}$ --- we compute $1.22$, $0.43$, and $0.03\,\deg$
for \mi , \mf , and \mm . These angle offsets are consistent with the values
given in \citet{2018arXiv180610564S} for the difference between the $z$-axis
as defined by the gas and stars.

Figure~\ref{fig:midplane} shows the relative $z$ location of the inferred midplane our imaginary observer
would determine as a function of azimuth for each galaxy, using their local ``solar neighborhood'' (the cylinder defined above). The 5--95 percentile range from the bootstrap procedure is shown as the dashed-line error bars. The $90\%$ interquartile
range for each galaxy is $185$, $162$, $84\,\pc$ for \mi , \mf , and \mm . In two of the three cases this varies by $\pm 100$pc depending on the azimuth along the Solar circle; in the third (\mm , which has the thinnest ``thin disk'' of stars, but the largest stellar mass) the variation is more like $\pm 50$ pc.

\begin{figure*}
\plotone{fig/midplane_fit.pdf}
\caption{The local midplane determined at the fiducial Solar radius
($8.2\,\kpc$) for the three FIRE galaxies \mi , \mf , and \mm\ ({\em left},
{\em center}, and {\em right} panels). The local midplane is determined at a
position $\phi$ by taking the median height of all stars within $R=0.5\,\kpc$
and $z=1\,\kpc$ (in cylindrical coordinates). The procedure is performed again
using the new height $10$ times to converge on the local midplane height. In
order to allow for the possibility that the fiducial Galactocentric coordinate
system is incorrect, we subtract the best fit $A\sin{(\phi+B)}+C$ curve from
each panel --- this figure is reproduced with the original midplane
determination (i.e. before subtracting the best fit sine curve) in
Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on all stars within
a $2\,\kpc$ height of the fiducial midplane to determine error bars (95\uth
and 5\uth percentiles), which we report as dashed lines.}
\label{fig:midplane}
\end{figure*}

\subsection{LSR Variations} \label{ssec:lsr_var}
We also expect
that the local standard of rest (LSR) should vary as a function of azimuth. We
perform this calculation in Appendix~\ref{app:lsr} to estimate , but performing a
best-fit subtraction to correct for misalignment of the original coordinate system (as in the previous section) is more involved. Since the variation in the LSR is
less pronounced than for the midplane, and since the dominant effect is limited to $J_r$ (Figure \ref{fig:many_orbit_wrong_ref}), we defer this calculation to future work.
